Lesson 3
========================================================

***

### What to Do First?
Notes: 

* Learn interesting things about the data  
* Produce summaries and visualizations for individual variables

***

### Pseudo-Facebook User Data
Notes:

* The data format needs to be examined carefully
* The data is generated by a model

```{r Pseudo-Facebook User Data}
getwd()
list.files()
pf <- read.csv("pseudo_facebook.tsv", sep = '\t')
names(pf)
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
#install.packages('ggplot2')
library(ggplot2)

qplot(x = dob_day, data = pf) +
    scale_x_discrete(breaks = 1:31)

# Get other themes
#install.packages('ggthemes', dependencies = TRUE)
library(ggthemes) 
theme_set(theme_minimal(18))

ggplot(aes(x = dob_day), data = pf) +
    geom_histogram() +
    scale_x_discrete(breaks = 1:31)
```

***

#### What are some things that you notice about this histogram?
Response: There is a disproportionately high number of 1's in the dataset.

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response: I don't use Facebook.

#### How many of your friends do you think saw that post?
Response: 0

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response: 0

***

### Perceived Audience Size
Notes:

* The preliminary result of underestimation in most cases is interesting.

***
### Faceting
Notes:

```{r Faceting}
qplot(x = dob_day, data = pf) +
    scale_x_discrete(breaks = 1:31) +
    facet_wrap(~ dob_month, ncol=3)
```

#### Let'۪s take another look at our plot. What stands out to you here?
Response: The excess 1's seem to be in January.

***

### Be Skeptical - Outliers and Anomalies
Notes:

* Usually we have to go in depth to say if an outlier is a real case or not

***

### Moira's Outlier
Notes:

* It's unlikely that someone has 10 million friends.

#### Which case do you think applies to Moira'۪s outlier?
Response: bad data about an extreme case

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(x = friend_count, data = pf)
```

#### How is this plot similar to Moira's first plot?
Response: It is similar because it is a distribution that looks like an exponential one with
some outliers as well.

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
qplot(x = friend_count, data = pf, xlim = c(0, 1000))

qplot(x = friend_count, data = pf) +
    scale_x_continuous(limits=c(0, 1000))
```

### Exploring with Bin Width
Notes:

* Changing the bin width can reveal interesting patterns in the data.

***

### Adjusting the Bin Width
Notes:

```{r}
qplot(x = friend_count, data = pf, binwidth = 25) +
    scale_x_continuous(limits=c(0, 1000))

ggplot(aes(x = friend_count), data = pf) +
    geom_histogram(binwidth = 25) +
    scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))
```


### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
    scale_x_continuous(limits = c(0, 1000),
                       breaks = seq(0, 1000, 50)) +
    facet_wrap(~ gender)
```

***

### Omitting NA Values
Notes:

* NA is an encoding for missing values

```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
    scale_x_continuous(limits = c(0, 1000),
                       breaks = seq(0, 1000, 50)) +
    facet_wrap(~ gender)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response: women

#### What's the difference between the median friend count for women and men?
Response: 22

#### Why would the median be a better measure than the mean?
Response: Because the distributions are very skewed.

***

### Tenure
Notes:

```{r Tenure}
ggplot(aes(x = tenure), data = pf) +
    geom_histogram(binwidth = 30, color = 'black', fill = '#099DD9')
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
ggplot(aes(x = tenure / 365), data = pf) +
    scale_x_continuous(limits = c(0, 8),
                       breaks = seq(0, 8, 1)) +
    geom_histogram(binwidth = 0.25, color = 'black', fill = '#F79420')
```

***

### Labeling Plots
Notes:

* The plots should speak for themselves

```{r Labeling Plots}
ggplot(aes(x = tenure / 365), data = pf) +
    xlab("Number of years using Facebook") +
    ylab("Number of users in sample") +
    scale_x_continuous(limits = c(0, 8),
                       breaks = seq(0, 8, 1)) +
    geom_histogram(binwidth = 0.25, color = 'black', fill = '#F79420')
```

***

### User Ages
Notes:

```{r User Ages}
ggplot(aes(x = age), data = pf) +
    xlab("Age") +
    ylab("Number of users in sample") +
    scale_x_continuous(limits = c(10, 110),
                       breaks = seq(10, 110, 2)) +
    geom_histogram(binwidth = 2, color = 'black', fill = '#5760AB')
```

#### What do you notice?
Response: By far the most users are in the 18-20 years range.

***

### The Spread of Memes
Notes:

* The finding that memes never truly disappear is very interesting.

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

```{r}
qplot(x = friend_count, data = pf)

summary(pf$friend_count)

summary(log10(pf$friend_count + 1))

summary(sqrt(pf$friend_count + 1))

#install.packages('gridExtra')
library(gridExtra)

p1 <- ggplot(aes(x = friend_count), data = pf) +
    xlab("Friend count") +
    ylab("Number of users in sample") +
    geom_histogram(binwidth = 100, color = 'black', fill = '#5760AB')
p2 <- ggplot(aes(x = friend_count), data = pf) +
    xlab("Friend count") +
    ylab("Number of users in sample") +
    scale_x_log10() +
    geom_histogram(binwidth = 0.2, color = 'black', fill = '#5760AB')
p3 <- ggplot(aes(x = friend_count), data = pf) +
    xlab("Friend count") +
    ylab("Number of users in sample") +
    scale_x_sqrt() +
    geom_histogram(binwidth = 2, color = 'black', fill = '#5760AB')

grid.arrange(p1, p2, p3)

```

***

### Add a Scaling Layer
Notes:

* scales can be added to existing plots(DRY)
* be careful with the axis names when adding scales

```{r Add a Scaling Layer}
countScale <- ggplot(aes(x = friend_count), data = pf) +
    geom_histogram() +
    scale_x_log10()

logScale <- qplot(x = log10(friend_count), data = pf)

grid.arrange(countScale, logScale, ncol = 2)
```

***


### Frequency Polygons

```{r Frequency Polygons}

qplot(x = friend_count,
      data = subset(pf, !is.na(gender)),
      binwidth = 10) +
    scale_x_continuous(limits = c(0, 1000),
                       breaks = seq(0, 1000, 50)) +
    facet_wrap(~ gender)

qplot(x = friend_count, y = ..count../sum(..count..),
      data = subset(pf, !is.na(gender)),
      xlab = "Friend Count",
      ylab = "Proportion of Users with that friend count",
      geom = "freqpoly", binwidth = 10, color = gender) +
    scale_x_continuous(limits = c(0, 1000),
                       breaks = seq(0, 1000, 50))

ggplot(aes(x = friend_count, y = ..count../sum(..count..)), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender)) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  xlab('Friend Count') +
  ylab('Percentage of users with that friend count')

```

***

### Likes on the Web
Notes:

* the ticks on the log scale are from the normal plot

```{r Likes on the Web}

ggplot(aes(x = www_likes), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender)) +
  scale_x_log10() +
  xlab('Likes Count') +
  ylab('Count of users with that like count')

ggplot(aes(x = log10(www_likes)), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender)) +
  scale_x_continuous()+
  xlab('Likes Count') +
  ylab('Count of users with that like count')

ggplot(aes(x = www_likes, y = ..count../sum(..count..)), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender), binwidth = 1000) +
  scale_x_continuous() +
  xlab('Likes Count') +
  ylab('Percentage of users with that like count')

```


***

### Box Plots
Notes:

```{r Box Plots}
by(pf$www_likes, pf$gender, sum)


```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}

```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}

```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}

```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!