Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv("../lesson3/pseudo_facebook.tsv", sep = "\t")

ggplot(pf, aes(x = age, y = friend_count)) +
    geom_point()
```

***

#### What are some things that you notice right away?
Response:

* Most high values are concentrated in the 18-28 age range
* there are some unusually high values in the >90 age range

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(pf, aes(x = age, y = friend_count)) +
    geom_point(alpha = 1/20) +
    xlim(13, 90)
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(pf, aes(x = age, y = friend_count)) +
    geom_jitter(alpha = 1/20) +
    xlim(13, 90)
```

#### What do you notice in the plot?
Response:

We can more easily see where the bulk of the data is located and how the friend
count is dependent of the age.

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(pf, aes(x = age, y = friend_count + 1)) +
    geom_jitter(alpha = 1/20) +
    xlim(13, 90) +
    coord_trans(y = "sqrt")
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(pf, aes(x = age, y = friend_count)) +
    geom_point(alpha = 1/20, position = position_jitter(h = 0)) +
    xlim(13, 90) +
    coord_trans(y = "sqrt")
```

#### What do you notice?

In the previous plots the points seemed to be very close to the Ox line because to the scales. Now we see that especially for the middle age group, there is a gap between y= 0 and the mean.

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(pf, aes(x = age, y = friendships_initiated)) +
    geom_point(alpha = 1/20, position = position_jitter(h = 0)) +
    xlim(13, 90) +
    coord_trans(y = "sqrt")
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(dplyr)

pf.fc_by_age <- pf %>%
    group_by(age) %>%
    summarise(friend_count_mean = mean(friend_count),
              friend_count_median = median(friend_count) * 1.0,
              n = n()) %>%
    arrange(age)

pf.fc_by_age
```

Create your plot!

```{r Conditional Means Plot}
ggplot(pf.fc_by_age, aes(x = age, y = friend_count_mean)) + geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(pf, aes(x = age, y = friend_count)) +
    geom_point(alpha = 1/20, position = position_jitter(h = 0), color = "orange") +
    xlim(13, 90) +
    coord_trans(y = "sqrt") +
    geom_line(stat = "summary", fun.y = mean) +
    geom_line(stat = "summary", fun.y = quantile, probs = .1, linetype = 2, color = "blue") +
    geom_line(stat = "summary", fun.y = quantile, probs = .5, linetype = 1, color = "blue") +
    geom_line(stat = "summary", fun.y = quantile, probs = .9, linetype = 2, color = "blue") 
```

#### What are some of your observations of the plot?
Response:

* the mean is larger than the median for every age
***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age, pf$friend_count)
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age <= 70) ,cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

* spearman correlation is more robust to outliers 

***

## Create Scatterplots
Notes:

```{r}
ggplot(pf, aes(x = www_likes_received, y = likes_received)) +
    xlim(0, quantile(pf$www_likes_received, 0.95)) +
    ylim(0, quantile(pf$likes_received, 0.95)) +
    geom_point(alpha = 1/20, position = position_jitter(h = 0))
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(pf, aes(x = www_likes_received, y = likes_received)) +
    xlim(0, quantile(pf$www_likes_received, 0.95)) +
    ylim(0, quantile(pf$likes_received, 0.95)) +
    geom_point(alpha = 1/20, position = position_jitter(h = 0)) +
    geom_smooth(method = "lm", color = "red", width = 2)
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
with(pf, cor(www_likes_received, likes_received))
```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
# install.packages('alr3')
library(alr3)
```

Create your plot!

```{r Temp vs Month}
data(Mitchell)
str(Mitchell)

ggplot(Mitchell, aes(x = Month, y = Temp)) +
    geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
with(Mitchell, cor(Temp, Month))
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(Mitchell, aes(x = Month, y = Temp)) +
    geom_point() +
    scale_x_discrete(breaks = seq(0, max(Mitchell$Month), 12))
```

***

### A New Perspective

What do you notice?
Response: There seems to be periodicity in the data.

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
pf$age_with_months <- pf$age + (12 - pf$dob_month) / 12
```

***

### Age with Months Means

```{r Age with Months Means}

```

Programming Assignment
```{r Programming Assignment}
pf.fc_by_age_months <- pf %>%
    group_by(age_with_months) %>%
    summarise(friend_count_mean = mean(friend_count),
              friend_count_median = median(friend_count) * 1,
              n = n()) %>%
    arrange(age_with_months)
pf.fc_by_age_months
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(pf.fc_by_age_months, aes(x = age_with_months, y = friend_count_mean)) +
    xlim(13, 71) +
    ylim(0, 500) +
    geom_line()
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
ggplot(pf.fc_by_age_months, aes(x = age_with_months, y = friend_count_mean)) +
    xlim(13, 71) +
    ylim(0, 500) +
    geom_point(alpha = 1/3) +
    geom_smooth()
```

***

### Which Plot to Choose?
Notes:

* It depends on the level of detail we want

***

### Analyzing Two Variables
Reflection:

* It is always a good idea to begin with a scatterplot.
* When analyzing two variables models might lose information. 
* Correlation is a measure of the linear relationship and ignores things like periodicity.

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

